---
title: "Getting Started"
author: "Matthew Mark Strasiotto"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(pipetree)
```

# Intoduction to Pipetree

The goal of pipetree is to make navigating the project structure of the 
speed-extract project as simple as possible, and to automate the most commonly
used tasks in accessing and exploring the dataset.

The most important functionalities provided are as follows:

1. Consistent access of a given dataset's root directory accross remote
   volumes, whose mount points may change.
1. Consistent Access to the filetree structure, that is used by the any SPEED-EXTRACT
   pipeline workflow/dataset.
1. Straightforward access to the [drake::drake_cache()] for a particular dataset.
   
   IE- Immediate access to the results, outputs, and targets generated by the data pipeline,
   as soon as these results are available.
